---
title: "Comparison between prior-exp-test experiment and exp-test experiment"
author: "Sebastian Schuster"
date: "10/27/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, include=FALSE}
library(ramify)
library(ggplot2)
library(dplyr)
library(tidyr)
library(data.table)
library(gridExtra)
library(DescTools)
library(splines)
library(stringr)
source("helpers.R")



theme_set(theme_bw())

colscale = scale_color_manual(values=c("#7CB637", "#666666", "#4381C1")) 
colscale_fill = scale_fill_manual(values=c("#7CB637", "#666666", "#4381C1")) 

remove_quotes = function(d) {
  if (!is.null(d$modal)) {
    d$modal = gsub('"', '', d$modal)
  }
  d$color = gsub('"', '', d$color)
  d$speaker_type = gsub('"', '', d$speaker_type)

  return(d)
}

format_data = function(d) {
  
    drops <- c("modal1","rating1")
  d2 = d[ , !(names(d) %in% drops)]
  setnames(d2, old=c("rating2","modal2"), new=c("rating", "modal"))

  drops <- c("modal2","rating2")
  d3 = d[ , !(names(d) %in% drops)]
  setnames(d3, old=c("rating1","modal1"), new=c("rating", "modal"))

  drops <- c("modal2", "rating2", "modal1", "rating1")
  d4 = d[ , !(names(d) %in% drops)]
  d4$rating = d4$rating_other
  d4$modal = "other"

  d = rbind(d2, d3, d4)

  d$modal = gsub('"', '', d$modal)
  d$modal = factor(d$modal, levels = c("might", "probably", "other"), ordered = T)
  
  d$percentage_blue_f = factor(d$percentage_blue)
  d_blue = d %>% filter(., grepl("blue", sentence2))
  d_orange = d %>% filter(., grepl("orange", sentence2))

  d_orange_reverse = d_orange
  d_orange_reverse$percentage_blue = 100-d_orange$percentage_blue

  d_comparison = rbind(d_blue, d_orange_reverse)
  d_comparison$blue= grepl("blue", d_comparison$sentence2)
  d_comparison$percentage_blue_f = factor(d_comparison$percentage_blue)

  return(d_comparison)
  
}


```

```{r data, echo=FALSE}
# Load data

# Experiment 1

trials.exp1 = remove_quotes(format_data(read.csv("../../1_talker_specific_adaptation_prior_exp_test/data/1_talker_specific_adaptation_prior_exp_test-trials.csv")))
exp_trials.exp1 = remove_quotes(read.csv("../../1_talker_specific_adaptation_prior_exp_test/data/1_talker_specific_adaptation_prior_exp_test-exp_trials.csv"))

conditions.exp1 = read.csv("../../1_talker_specific_adaptation_prior_exp_test/data/1_talker_specific_adaptation_prior_exp_test-condition.csv")

trials.exp1 = merge(trials.exp1, conditions.exp1 %>% select(workerid, condition), by=c("workerid"))

subj_info.exp1 = read.csv("../../1_talker_specific_adaptation_prior_exp_test/data/1_talker_specific_adaptation_prior_exp_test-subject_information.csv") %>% select(workerid, noticed_manipulation)

trials.exp1 = merge(trials.exp1, subj_info.exp1, by=c("workerid"))


trials.exp1[, c("test_order", "first_speaker_type", "second_speaker_type", "confident_speaker")] = str_split(trials.exp1$condition, "_", simplify=T)


# Experiment 2
trials.exp2 = remove_quotes(format_data(read.csv("../data/2_talker_specific_adaptation_replication-trials.csv")))
exp_trials.exp2 = remove_quotes(read.csv("../data/2_talker_specific_adaptation_replication-exp_trials.csv"))

conditions.exp2 = read.csv("../data/2_talker_specific_adaptation_replication-condition.csv")

trials.exp2 = merge(trials.exp2, conditions.exp2 %>% select(workerid, condition), by=c("workerid"))

subj_info.exp2 = read.csv("../data/2_talker_specific_adaptation_replication-subject_information.csv") %>% select(workerid, noticed_manipulation)

trials.exp2 = merge(trials.exp2, subj_info.exp2, by=c("workerid"))


trials.exp2[, c("test_order", "first_speaker_type", "second_speaker_type", "confident_speaker")] = str_split(trials.exp2$condition, "_", simplify=T)

```

## Catch trial performance

```{r catch_trials, echo=FALSE}

get_correct_catch_trial_counts = function (data) {
  ret = data %>% 
  filter(., catch_trial == 1) %>%
  group_by(workerid) %>%
  summarise(catch_perf = sum(catch_trial_answer_correct))
  
  return(ret)
}

EXCLUDE_BELOW = 9

catch_trial_perf.exp1 = get_correct_catch_trial_counts(exp_trials.exp1)

exclude = catch_trial_perf.exp1 %>%
  filter(catch_perf < EXCLUDE_BELOW) %>%
  .$workerid


print(paste("Excluded", length(exclude), "participants based on catch-trial performance from Experiment 1."))

d.exp1 = trials.exp1 %>% filter(., !(workerid %in% exclude))



catch_trial_perf.exp2 = get_correct_catch_trial_counts(exp_trials.exp2)

exclude = catch_trial_perf.exp2 %>%
  filter(catch_perf < EXCLUDE_BELOW) %>%
  .$workerid


print(paste("Excluded", length(exclude), "participants based on catch-trial performance from Experiment 2."))


#final data
d.exp2 = trials.exp2 %>% filter(., !(workerid %in% exclude))

```



## Exclusion of random guesses

We further exclude participants who seem to provide random ratings independent of the scene that they are seeing. We quantify this by computing the mean rating for each utterance across all trials for each participant and computing the correlation between a participant's actual ratings and their mean rating. A high correlation is unexpected and indicates that a participant chose ratings at random. We therefore also exclude the data from participants for whom this correlation is larger than 0.75.

```{r random_guess_exclusions, echo=FALSE}

d_overall_means = d.exp1 %>%
  group_by(modal, workerid) %>% 
  summarise(rating_m_overall = mean(rating))

d_indiv_means =  d.exp1 %>%
  group_by(modal,percentage_blue, workerid) %>% 
  summarise(rating_m = mean(rating))


d_indiv_merged = merge(d_indiv_means, d_overall_means, by=c("workerid", "modal"))

cors = d_indiv_merged %>%
  group_by(workerid) %>%
  summarise(corr = cor(rating_m, rating_m_overall))


exclude = cors %>%
  filter(corr > 0.75) %>%
  .$workerid


print(paste("Excluded", length(exclude), "participants based on random responses in Experiment 1."))

d.exp1 = d.exp1 %>% filter(!(workerid %in% exclude))



d_overall_means = d.exp2 %>%
  group_by(modal, workerid) %>% 
  summarise(rating_m_overall = mean(rating))

d_indiv_means =  d.exp2 %>%
  group_by(modal,percentage_blue, workerid) %>% 
  summarise(rating_m = mean(rating))


d_indiv_merged = merge(d_indiv_means, d_overall_means, by=c("workerid", "modal"))

cors = d_indiv_merged %>%
  group_by(workerid) %>%
  summarise(corr = cor(rating_m, rating_m_overall))


exclude = cors %>%
  filter(corr > 0.75) %>%
  .$workerid


print(paste("Excluded", length(exclude), "participants based on random responses in Experiment 2."))

d.exp2 = d.exp2 %>% filter(!(workerid %in% exclude))


d.exp1$experiment = "prior-exposure-test"
d.exp2$experiment = "exposure-test"

d = rbind(d.exp1, d.exp2)

```


## Comparisons plots


```{r aggregate_plots, echo=FALSE, fig.width=10, fig.height=8}

d_means = d %>%
  group_by(workerid, percentage_blue, modal, speaker_type, experiment) %>% 
  summarise(participant_mean = mean(rating)) %>%
  group_by(percentage_blue, modal, speaker_type, experiment) %>%
  summarise(mu = mean(participant_mean),
         ci_high = ci.high(participant_mean), 
         ci_low = ci.low(participant_mean))

d_means %>% filter(speaker_type != "prior") %>%
  ggplot(aes(x=percentage_blue, y=mu, col=experiment)) + 
    xlab("% blue gumballs") +
    ylab("mean ratings") +
    geom_errorbar(aes(ymin=mu-ci_low, ymax=mu+ci_high), width=.1) +
    geom_line() +
    geom_point(size=1) +
    guides(col=guide_legend(title="Expr.")) +
    theme(legend.position="bottom") + 
    facet_wrap(~speaker_type+modal, nrow = 2)

```

