---
title: "Prior-exposure-test Experiment (all data combined)"
author: "Sebastian Schuster"
date: "10/24/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, include=FALSE}
library(ramify)
library(ggplot2)
library(dplyr)
library(tidyr)
library(data.table)
library(gridExtra)
library(DescTools)
library(splines)
library(stringr)
library(lme4)
library(lmerTest)
source("helpers.R")



theme_set(theme_bw())

colscale = scale_color_manual(values=c("#7CB637", "#666666", "#4381C1")) 
colscale_fill = scale_fill_manual(values=c("#7CB637", "#666666", "#4381C1")) 

remove_quotes = function(d) {
  if (!is.null(d$modal)) {
    d$modal = gsub('"', '', d$modal)
  }
  d$color = gsub('"', '', d$color)
  d$speaker_type = gsub('"', '', d$speaker_type)

  return(d)
}

format_data = function(d) {
  
    drops <- c("modal1","rating1")
  d2 = d[ , !(names(d) %in% drops)]
  setnames(d2, old=c("rating2","modal2"), new=c("rating", "modal"))

  drops <- c("modal2","rating2")
  d3 = d[ , !(names(d) %in% drops)]
  setnames(d3, old=c("rating1","modal1"), new=c("rating", "modal"))

  drops <- c("modal2", "rating2", "modal1", "rating1")
  d4 = d[ , !(names(d) %in% drops)]
  d4$rating = d4$rating_other
  d4$modal = "other"

  d = rbind(d2, d3, d4)

  d$modal = gsub('"', '', d$modal)
  d$modal = factor(d$modal, levels = c("might", "probably", "other"), ordered = T)
  
  d$percentage_blue_f = factor(d$percentage_blue)
  d_blue = d %>% filter(., grepl("blue", sentence2))
  d_orange = d %>% filter(., grepl("orange", sentence2))

  d_orange_reverse = d_orange
  d_orange_reverse$percentage_blue = 100-d_orange$percentage_blue

  d_comparison = rbind(d_blue, d_orange_reverse)
  d_comparison$blue= grepl("blue", d_comparison$sentence2)
  d_comparison$percentage_blue_f = factor(d_comparison$percentage_blue)

  return(d_comparison)
  
}


```

```{r data, echo=FALSE}
# Load data

trials = remove_quotes(format_data(read.csv("../data/3_talker_specific_adaptation_prior_exp_test_id-trials.csv")))
exp_trials = remove_quotes(read.csv("../data/3_talker_specific_adaptation_prior_exp_test_id-exp_trials.csv"))

conditions = read.csv("../data/3_talker_specific_adaptation_prior_exp_test_id-condition.csv")

subj_info = read.csv("../data/3_talker_specific_adaptation_prior_exp_test_id-subject_information.csv") %>% select(workerid, noticed_manipulation)

trials = merge(trials, conditions %>% select(workerid, condition), by=c("workerid"))

trials = merge(trials, subj_info, by=c("workerid"))

trials[, c("test_order", "first_speaker_type", "second_speaker_type", "confident_speaker")] = str_split(trials$condition, "_", simplify=T)

```

## Catch trial performance

```{r catch_trials, echo=FALSE}

get_correct_catch_trial_counts = function (data) {
  ret = data %>% 
  filter(catch_trial == 1) %>%
  group_by(workerid) %>%
  summarise(catch_perf = sum(catch_trial_answer_correct), catch_prop = sum(catch_trial_answer_correct)/n())
  
  return(ret)
}

EXCLUDE_BELOW = 11

catch_trial_perf = get_correct_catch_trial_counts(exp_trials)

exclude = catch_trial_perf %>%
  filter(catch_perf < EXCLUDE_BELOW) %>%
  .$workerid


print(paste("Excluded", length(exclude), "participants based on catch-trial performance."))


#final data
d.exp3 = trials %>% filter(., !(workerid %in% exclude))

```

```{r perf_plot_1, fig.height=2,fig.width=10, echo=FALSE}

cutoff <- data.frame(yintercept=EXCLUDE_BELOW-0.5, cutoff=factor(EXCLUDE_BELOW-0.5))


ggplot(catch_trial_perf, aes(x=factor(workerid), y=catch_perf)) + geom_bar(stat = "identity") + ggtitle("Catch trial performace") + xlab ("participant") + ylab("correct responses") + geom_hline(aes(yintercept=yintercept, linetype=cutoff, color="red"), data=cutoff) + theme(legend.position="none")

```

## Load data from Exp 1.

```{r load_exp_1_data}

trials = remove_quotes(format_data(read.csv("../../1_talker_specific_adaptation_prior_exp_test/data/1_talker_specific_adaptation_prior_exp_test-trials.csv")))
exp_trials = remove_quotes(read.csv("../../1_talker_specific_adaptation_prior_exp_test/data/1_talker_specific_adaptation_prior_exp_test-exp_trials.csv"))

conditions = read.csv("../../1_talker_specific_adaptation_prior_exp_test/data/1_talker_specific_adaptation_prior_exp_test-condition.csv")

subj_info = read.csv("../../1_talker_specific_adaptation_prior_exp_test/data/1_talker_specific_adaptation_prior_exp_test-subject_information.csv") %>% select(workerid, noticed_manipulation)

trials = merge(trials, conditions %>% select(workerid, condition), by=c("workerid"))

trials = merge(trials, subj_info, by=c("workerid"))

trials[, c("test_order", "first_speaker_type", "second_speaker_type", "confident_speaker")] = str_split(trials$condition, "_", simplify=T)



EXCLUDE_BELOW = 0.75

catch_trial_perf = get_correct_catch_trial_counts(exp_trials)



exclude = catch_trial_perf %>%
  filter(catch_prop < EXCLUDE_BELOW) %>%
  .$workerid


print(paste("Excluded", length(exclude), "participants based on catch-trial performance in Exp 1.."))


#final data
d.exp1 = trials %>% filter(., !(workerid %in% exclude))

#combine data from exp1 and exp3

d = rbind(d.exp1, d.exp3)

```




## Aggregated results


```{r aggregate_plots, echo=FALSE, fig.width=10, fig.height=5}



plot_conditions = function(d, plot_title) {
  
  
  d_means = d %>%
  group_by(workerid, percentage_blue, modal, speaker_type) %>% 
  summarise(participant_mean = mean(rating)) %>%
  group_by(percentage_blue, modal, speaker_type) %>%
  summarise(mu = mean(participant_mean),
         ci_high = ci.high(participant_mean), 
         ci_low = ci.low(participant_mean))

  
  p1 = ggplot(d_means, aes(x=percentage_blue, y=mu, col=modal)) + 
  xlab("% blue gumballs") +
  ylab("mean ratings") +
  geom_errorbar(aes(ymin=mu-ci_low, ymax=mu+ci_high), width=.1) +
  geom_line() +
  geom_point(size=1) +
  guides(col=guide_legend(title="Expr.")) +
  colscale +
  theme(legend.position="bottom") + 
  ggtitle(plot_title) +
  facet_wrap(~speaker_type) +
  geom_vline(xintercept=60, lty=3, col="gray", size=1)

  p2 = ggplot(d, aes(x=percentage_blue_f, y=rating, fill=modal)) + 
  geom_boxplot() +
  ggtitle(plot_title) + xlab("percentage") + colscale_fill +   theme(legend.position="bottom")

  by_participant = ggplot( d %>% filter(speaker_type != "prior") %>%
  group_by(workerid, percentage_blue, modal, speaker_type) %>% 
  summarise(participant_mean = mean(rating)) , 
    aes(x=percentage_blue, y=participant_mean, col=modal, lty=speaker_type)
  ) + 
  xlab("% blue gumballs") +
  ylab("mean ratings") +
  geom_line() +
  geom_point(size=1) +
  guides(col=guide_legend(title="Expr.")) +
  colscale +
  theme(legend.position="bottom") + ggtitle(plot_title) + facet_wrap(~workerid, ncol = 8)
  return(list("p1" = p1, "p2" = p2, "by_participant" = by_participant))

}



  ps1 = plot_conditions(d, "")

  plot(ps1$p1)
```


## Comparison across conditions

```{r comparison_plots, echo=FALSE, fig.width=10, fig.height=5}


  d_means = d %>%
  group_by(workerid, percentage_blue, modal, speaker_type) %>% 
  summarise(participant_mean = mean(rating)) %>%
  group_by(percentage_blue, modal, speaker_type) %>%
  summarise(mu = mean(participant_mean),
         ci_high = ci.high(participant_mean), 
         ci_low = ci.low(participant_mean))

  
  
  d_means %>% 
    ggplot(aes(x=percentage_blue, y=mu, col=speaker_type)) + 
      xlab("% blue gumballs") +
      ylab("mean ratings") +
      geom_errorbar(aes(ymin=mu-ci_low, ymax=mu+ci_high), width=.1) +
      geom_line() +
      geom_point(size=1) +
      guides(col=guide_legend(title="Spoeaker type:")) +
      theme(legend.position="bottom") + 
      facet_wrap(~modal)

```

## Individual responses

```{r indiv_plots, echo=FALSE, fig.width=10, fig.height=20}

    
  plot(ps1$by_participant)


```

## AUC computation

We use the `AUC` function with the `splines` method to directly compute the AUC.

```{r auc_method_1, echo=FALSE, warning=FALSE, message=FALSE}

auc_method2 = function(d) {
  auc = AUC(x=d$percentage_blue, y=d$rating_m, method="spline")
  return(auc)
}


auc_for_participants = function(d, method) {
  
  d[d$color=="orange",]$percentage_blue = 100 - d[d$color=="orange",]$percentage_blue 
  
  aucs = d %>% 
    group_by(workerid) %>% 
    summarize(.groups = "drop_last", test_order = first(test_order),
              first_speaker_type = first(first_speaker_type),
              confident_speaker = first(confident_speaker))
    
    
  aucs$auc_might = 0
  aucs$auc_probably = 0

  
  i = 1
  
  for (wid in unique(d$workerid)) {
    d.might_ratings = d %>% 
      filter (workerid == wid) %>%
      filter (modal == "might") %>%
      group_by(workerid, percentage_blue) %>%
      summarise(rating_m = mean(rating))
    
    aucs$auc_might[i] = method(d.might_ratings)

    d.probably_ratings = d %>% 
      filter (workerid == wid) %>%
      filter (modal == "probably") %>%
      group_by(workerid, percentage_blue) %>%
      summarise(rating_m = mean(rating))
    
    aucs$auc_probably[i] = method(d.probably_ratings)

    i = i + 1
  }
  
  aucs$auc_diff = aucs$auc_might - aucs$auc_probably
  
  return(aucs)
}

#AUCs for cautious speaker ratings 
aucs.cautious = d %>% filter(speaker_type == "cautious") %>% auc_for_participants(., method=auc_method2)

#AUCs for confident speaker ratings 
aucs.confident = d %>% filter(speaker_type == "confident") %>% auc_for_participants(., method=auc_method2)



```


```{r auc_method1_plots, fig.width=4, fig.height=4, echo=FALSE}

aucs.cautious$cond = "cautious (might-biased)"

aucs.confident$cond = "confident (probably-biased)"


aucs.all = rbind(aucs.cautious, aucs.confident)


aucs.all = aucs.all %>% 
  group_by(., cond) %>% 
  summarise(., auc_diff_m = mean(auc_diff), 
               ci_high = ci.high(auc_diff), 
               ci_low = ci.low(auc_diff))




ggplot(aucs.all, aes(x=0, y=auc_diff_m, color=cond)) +
    geom_errorbar(aes(ymin=auc_diff_m-ci_low, ymax=auc_diff_m+ci_high), width=.1) +
    geom_point() +
    xlab("") +
    ylab("AUC difference (might ratings - probably ratings)") +
    theme(axis.ticks=element_blank(), axis.text.x=element_blank(),
          panel.grid.minor=element_blank(), 
          plot.background=element_blank()) +
    xlim(-.2, .2) 

```

t-test and regression model with control variables:

```{r t_test_method1, echo=FALSE}

print(t.test(aucs.cautious$auc_diff, aucs.confident$auc_diff, var.equal=TRUE))

auc_d = rbind(aucs.cautious, aucs.confident)

auc_d$cond = factor(auc_d$cond)
auc_d$test_order = factor(auc_d$test_order)
auc_d$first_speaker_type = factor(auc_d$first_speaker_type)
auc_d$confident_speaker = factor(auc_d$confident_speaker)

contrasts(auc_d$cond) = contr.sum(2)
contrasts(auc_d$test_order) = contr.sum(2)
contrasts(auc_d$first_speaker_type) = contr.sum(2)
contrasts(auc_d$confident_speaker) = contr.sum(2)

model = lmer(formula= auc_diff ~ cond + test_order + first_speaker_type + confident_speaker + first_speaker_type * cond + (1| workerid), data = auc_d)

print(summary(model))

```

### Clustering analyses

```{r clustering_1}

library(mclust)
aucs_diff = merge(aucs.cautious, aucs.confident, by=c("workerid"))
aucs_diff$diff_of_diffs = aucs_diff$auc_diff.x - aucs_diff$auc_diff.y

aucs_diff %>% ggplot(aes(x=diff_of_diffs)) + geom_density() + geom_jitter(aes(y=0), width=0, height=0.001)  + ggtitle("Raw data + estimated density")


```

### Gaussian mixture models of diffeences of AUC differences

1 Cluster

```{r cluster_mclust1}

fit1 = Mclust(aucs_diff$diff_of_diffs, G=1)
print(summary(fit1, parameters=2))

```


2 Clusters

```{r cluster_mclust2}

fit2 = Mclust(aucs_diff$diff_of_diffs, G=2)
print(summary(fit2, parameters=T))

```


3 Clusters

```{r cluster_mclust3}

fit3 = Mclust(aucs_diff$diff_of_diffs, G=3)
print(summary(fit3, parameters=T))

```

According to the Bayesian information criterion, a model with two clusters describes the data best.

Fitted model:

```{r }
aucs_diff %>% 
  ggplot(aes(x=diff_of_diffs)) + 
    geom_jitter(aes(y=0, color=first_speaker_type.x), width=0, height=0.001)  +
    ggtitle("Raw data + Components of gaussian mixture") + 
    stat_function(fun = dnorm, args = list(mean = fit2$parameters$mean[1], sd = sqrt(fit2$parameters$variance$sigmasq[1]))) + 
    stat_function(fun = dnorm, args = list(mean = fit2$parameters$mean[2], sd = sqrt(fit2$parameters$variance$sigmasq[2])))
```

## Compute likelihoods based on the adaptation model

```{r analyze_likelihoods, echo=F, message=F}

# load model estimates (in original adaptation repo)
d_likelihoods.cautious.exp3 = read.csv("~/Dropbox/Uni/RA/adaptation/adaptation/models/4_individual_differences/results/id-03_prior_exp_post_id-cautious.likelihood")

d_likelihoods.cautious.exp1 = read.csv("~/Dropbox/Uni/RA/adaptation/adaptation/models/4_individual_differences/results/id-01__prior_exp_post-cautious.likelihoods")
  
d_likelihoods.cautious = rbind(d_likelihoods.cautious.exp1, d_likelihoods.cautious.exp3)

d_likelihoods.confident.exp3 = read.csv("~/Dropbox/Uni/RA/adaptation/adaptation/models/4_individual_differences/results/id-03_prior_exp_post_id-confident.likelihood")

d_likelihoods.confident.exp1 = read.csv("~/Dropbox/Uni/RA/adaptation/adaptation/models/4_individual_differences/results/id-01__prior_exp_post-confident.likelihoods")


d_likelihoods.confident = rbind(d_likelihoods.confident.exp1, d_likelihoods.confident.exp3)
  
d_likelihoods.merged = merge(d_likelihoods.cautious, 
                               d_likelihoods.confident,
                               by=c("workerid", "condition"))
  
d_likelihoods.merged = d_likelihoods.merged %>% 
  mutate(most_likely_model = 
           case_when(likelihood.x > likelihood.y ~ "cautious", 
                     TRUE ~ "confident" ) ) %>%
  rename(likelihood.cautious = likelihood.x, 
         likelihood.confident = likelihood.y) 
  
# compute likelihood ratio
d.lr = d_likelihoods.merged %>% 
  pivot_longer(starts_with("likelihood."), 
               names_prefix="likelihood.", 
               names_to="likelihood") %>% 
  group_by(workerid, condition, most_likely_model) %>%
  summarise(likelihood_ratio = diff(value)) %>%
  ungroup() %>%
  select(workerid, condition, likelihood_ratio)
  
d_likelihoods.merged =  merge(d_likelihoods.merged, 
                              d.lr, 
                              by=c("workerid", "condition"))
  
d.all = merge(d_likelihoods.merged, d %>%
                  select(-condition) %>% 
                  group_by(workerid, 
                           speaker_type, 
                           test_order, 
                           noticed_manipulation) %>% 
                  summarise(
                    first_speaker_type = first(first_speaker_type)) %>%
                  rename(condition = speaker_type), 
                by=c("workerid", "condition"))


# plot inferred model by condition  
d.all %>% 
  ggplot(aes(x=most_likely_model, fill=first_speaker_type)) + 
    geom_bar(position = "dodge") + 
    facet_wrap(~condition)
  
d.all %>% 
  ggplot(aes(x=most_likely_model)) + 
    geom_bar()
  
d.aggregated_stats = d.all %>% 
  filter(condition != "prior") %>%
    group_by(workerid, 
             first_speaker_type, 
             test_order, 
             noticed_manipulation) %>%
    summarise(cautious_count = sum(most_likely_model == "cautious"),
              confident_count = sum(most_likely_model == "confident"),
              aligned_count = sum(most_likely_model == condition)) %>% 
    mutate( first_adaptation_speaker_count = 
              case_when(first_speaker_type == "cautiousfirst" ~ cautious_count,
                        TRUE ~ confident_count)) 
  


d.aggregated_stats %>% 
  ggplot(aes(x=first_adaptation_speaker_count)) +
    geom_bar() +
    facet_grid(~first_speaker_type)
  
d.aggregated_stats %>% 
  ggplot(aes(x=aligned_count)) +
    geom_bar()
   
d.aggregated_stats %>% 
  ggplot(aes(x=aligned_count)) +
  geom_bar() + 
  facet_wrap(~noticed_manipulation)
 
d.post_test = d.all %>% 
  filter(condition != "prior") %>%
  mutate(condition = factor(condition), 
         test_order = factor(test_order), 
         first_speaker_type = factor(first_speaker_type),
         most_likely_model = factor(most_likely_model))

# use sum-coding    
contrasts(d.post_test$condition) = contr.sum(2)
colnames(contrasts(d.post_test$condition)) = c("cautious")
contrasts(d.post_test$test_order) = contr.sum(2)
colnames(contrasts(d.post_test$test_order)) = c("parallel")
contrasts(d.post_test$first_speaker_type) = contr.sum(2)
colnames(contrasts(d.post_test$first_speaker_type)) = c("cautious")

# logistic regression ME model predicting the most likely condition
model = glmer(formula = most_likely_model ~ condition + 
                test_order + 
                first_speaker_type + 
                first_speaker_type * condition + 
                (1| workerid), 
              data = d.post_test, family = "binomial")

print(summary(model))

# linear regression ME model predicting the likelihood ratio
model1 = lmer(formula=likelihood_ratio ~ 
               condition + 
               test_order + 
               first_speaker_type + 
               first_speaker_type * condition + 
               (1| workerid), 
             data = d.post_test)

print(summary(model1))


d.post_test = merge(d.post_test, d.all %>% 
                      filter(condition == "prior") %>% 
                      select(workerid, likelihood_ratio) %>% 
                      rename(prior_likelihood_ratio = likelihood_ratio))

# linear regression ME model predicting the likelihood ratio + predictor based on prior

model2 = lmer(formula= likelihood_ratio ~ condition + 
               test_order + 
               first_speaker_type + 
               prior_likelihood_ratio + 
               first_speaker_type * condition + 
               (1| workerid), 
             data = d.post_test)

print(summary(model2))

anova(model1, model2)

write.csv(d.post_test, "../../4_id_tasks/data/5_adaptation_data-merged.csv")


```      
        
```{r likelihood_plots,echo=FALSE, fig.width=10, fig.height=10}
      
      d_likelihoods.merged %>% 
        ggplot(aes(y=likelihood_ratio, 
                   x=(condition != "prior"), 
                   col=condition)) + 
          facet_wrap(~workerid) + geom_point() + geom_hline(yintercept=0, col="black", size=1)

```

### List of adapters:



```{r adapters_list, echo=F}
   knitr::kable(d.aggregated_stats %>% 
                  filter(aligned_count == 2))

  
  #prolific_ids = read.csv("../../proliferate/1_talker_specific_adaptation_prior_exp_test/1_talker_specific_adaptation_prior_exp_test-workerids.csv")

  # write them to file
  #d.aggregated_stats %>% 
  #  filter(aligned_count == 2) %>% 
  #  ungroup() %>%
  #  select(workerid) %>%
  #  merge(prolific_ids) %>% 
  #  select(workerid, prolific_participant_id) %>%
  #  write.csv(file="../../proliferate/1_talker_specific_adaptation_prior_exp_test/adapters-workerids.csv")

  #d.aggregated_stats %>% 
  #  filter(aligned_count == 1) %>% 
  #  ungroup() %>%
  #  select(workerid) %>%
  #  merge(prolific_ids) %>% 
  #  select(workerid, prolific_participant_id) %>%
  #  write.csv(file="../../proliferate/1_talker_specific_adaptation_prior_exp_test/nonadapters-workerids.csv")

    
```

### List of reverse adapters:

```{r reverseadapters_list, echo=F}
   knitr::kable(d.aggregated_stats %>% 
                  filter(aligned_count == 0))
```
